### 2015-01-06 response.txt
Checked through response with Yashar

We are grateful to the referee for careful reading of the 
manuscript and useful suggestions.

==================================================================================

The paper makes an interesting point that having extended sources 
can significantly increase the detection significance because such 
sources "occupy a larger area of the source plane, which may cover 
regions with higher magnifications".

I think it could be useful to see this point illustrated, perhaps 
with a magnification map and a curve showing magnification as a 
function of source size (at a given position).

Also, the paper argues that central images can offer good 
constraints on the core size and black hole mass.

It would be nice to have some intuition about which image 
properties provide the most information: position, brightness, 
resolved shape, or all of the above?

The text broaches this issue in Section 3.2 but does not go into 
detail. In both of these cases, my personal preference would be to 
have a more extended discussion (with figures) even if that means 
the paper becomes too long for ApJL. I do want to emphasize that 
this is a personal preference; the authors may be satisfied with 
the level of the discussion in its present form.

>> RESPONSE: We have added a figure (Figure 4) showing the magnification 
>> cross-section for two sources with different sizes. We have also 
>> added a few sentences to the text to explain this, as follows:

"This effect is illustrated in Figure 4 where the effects of source 
size on magnification are shown. The two top panels of Figure 4 
show the source plane magnification (flux magnification as a 
function of source position behind the lens) for two sources with 
different sizes (left: 100 pc, right: 500 pc in radius). For a 
compact source (top left panel) in the central regions the 
magnification approaches zero and the central image does not form. 
For more extended sources, however, even when the source is exactly 
behind the lens, the large extent of the source allows the covering of 
higher magnification regions, increasing the total flux 
magnification. The bottom panel shows the magnification as a 
function of source size, for a source that is located on the lens 
axis."

==================================================================================

My second comment is that there is little discussion of existing 
literature on gravitational lensing, central images, and 
supermassive black holes. Here are a few papers that seem worthy of 
mention (there may be others as well):

- Mao, Witt & Koopmans, http://adsabs.harvard.edu/abs/2001MNRAS.323..301M
- Bowman, Hewitt & Kiger, http://adsabs.harvard.edu/abs/2004ApJ...617...81B
- Rusin, Keeton & Winn, http://adsabs.harvard.edu/abs/2005ApJ...627L..93R
- Mao & Witt, http://adsabs.harvard.edu/abs/2012MNRAS.420..792M

>> RESPONSE: The citations and a short description have been added 
>> to the introduction.

==================================================================================

What was the source size used to generate Figure 1?

>> RESPONSE: The source has a Gaussian profile with FWHM of 1170 pc 
>> (rms of 500 pc). We have added this information to the caption:

" a background source with a Gaussian light profile (FWHM 
$\sim1100$ pc)..."

==================================================================================

How was the black hole mass chosen? I assume it follows the M-sigma 
relation, but that is not discussed explicitly (perhaps because the 
effective velocity dispersion of the model galaxy is not 
mentioned).

>> RESPONSE:

==================================================================================

The analysis uses circular mass distributions. How would it be 
affected by ellipticity and/or external shear?

>> We thank the referee for pointing out the ambiguity of the text 
>> about this. In fact, elliptical power-law models were used in 
>> the calculations and the deflection angles were computed using 
>> the FastEll code. We ran simulations with ellipticities of 0.2 
>> and 0.4. We have added a few words to clarify this, as follows:

"The stellar population is modeled as an elliptical cored 
power-law. The deflection angles for elliptical cored power-law 
models were calculated using FastEll (Barkana 1998)."

==================================================================================

"For double and quad image configurations, the theory of strong 
lensing predicts a third and a fifth image near the centers of 
lensing galaxies." Strictly speaking, that statement applies only 
under certain conditions. Even if those conditions are thought to 
apply to realistic galaxy models, it is important to understand 
that the theoretical statement is nuanced. For example, there is no 
central image if the surface mass density has a central cusp that 
is steeper than 1/r, or if the SMBH exceeds a threshold in mass. If 
the black hole mass is below that threshold, there can actually be 
more than one central image.

>> RESPONSE:

==================================================================================

"Unlike the other lensed images, which are magnified..." It is 
possible for non-central images to be demagnified. The only true 
guarantee from lens theory is that at least one image has a 
magnification greater than unity.

>> RESPONSE: We have modified the wording to "Realistic galaxy 
>> density profiles produce central images that are almost always 
>> demagnified, making their detection difficult."

==================================================================================

"The dashed curves in Figure 3 show the suppression of the fluxes 
of central images in presence of a 2e8 Msun SMBH." There is only 
one relevant dashed curve in Figure 3, and it shows a case with no 
black hole. The suppression effect is therefore seen not in the 
dashed curve itself, but in a comparison of the solid curves to the 
dashed curve.

>> RESPONSE: We have changed the wording to clarify this, as follows:

"A comparison of the thick black dashed curve to the solid black 
curve in Figure 3 shows the suppression of the fluxes of central 
images in presence of a 2e8 Msun SMBH."

==================================================================================

"The radio spectrum of candidate central images have been used" -- 
here the noun ("spectrum") is singular while the verb ("have") is 
plural.

>> DONE

==================================================================================

"generated maps with and without the the demagnified flux" -- 
double "the".

>> DONE

==================================================================================

"and nuisance parameters were later marginalize over" -- should be 
"marginalized".

>> DONE